Software Architectures to Integrate Workflow Engines in Science Gateways

### 2016-04-28 paper.tex
Fixed @merous note on description of pool architecture

\begin{figure*}
\centering
\def\svgwidth{1.5\columnwidth}
\input{figures/VIP.pdf_tex}
\caption{Architecture of the Virtual Imaging Platform (service
  invocation).  Step \texttt{3} corresponds to interaction \texttt{d}
  on Figure~\ref{fig:architectures}, step \texttt{2} corresponds to
  interaction \texttt{c$_1$}, and step \texttt{4} maps to interactions
  \texttt{b} and \texttt{c$_2$} (in VIP, workflow data transfers are
  performed by the tasks and embedded in their descriptions).}
\label{fig:vip-architecture}
\end{figure*}

@@

\subsection{Task encapsulation} \answerednote{Marc-e}{May I propose
  another name here? Either "Task-centric" or "Task Oriented". My
  reasoning is that the WE is just another task, as well as the
  sub-tasks themselves.}{Tristan}{I agree: sub-tasking is not the
  integration model, it is the process that engines use when they
  are integrated as tasks. I think 'task encapsulation' is more
  consistent  with 'service invocation' than 'task-oriented'}

@@

\begin{figure*}
\centering
\subfloat[CBRAIN-FSL integration.]{
\def\svgwidth{\columnwidth}
\input{figures/CBRAIN-FSL.pdf_tex}
\label{fig:cbrain-fsl-architecture}
} \hfill \subfloat[CBRAIN-PSOM integration. Figure
reproduced from~\cite{GLAT-16}]{ \def\svgwidth{\columnwidth}
  \input{figures/CBRAIN-PSOM.pdf_tex}
\label{fig:cbrain-psom-architecture}
}
\caption{CBRAIN architecture for workflow engine integration
  (task encapsulation).  1: User sends data and workflow execution request to
  storage server(s) and CBRAIN portal. 2: CBRAIN portal sends
  execution request to execution server on cluster. 3: Execution
  server transfers data from storage server(s). 4-5: Execution server
  starts workflow engine (FSL tool or PSOM master) via task
  scheduler. 6: Workflow engine submits sub-tasks to execution server
  (FSL tasks or PSOM agents). 7-8: Execution server starts sub-tasks
  through task scheduler. FSL sub-tasks will run locally
  instead of being submitted to CBRAIN through interaction 6. 8':
  (PSOM only) PSOM master and workers execute workflow. 9. Execution
  server transfers results to storage server(s). Interaction
  \texttt{b} in Figure~\ref{fig:architectures} is implemented by steps
  \texttt{4}, \texttt{5}, \texttt{7} and \texttt{8} (regular
  interactions with batch manager). Interaction \texttt{c} is
  implemented by steps \texttt{3} and \texttt{9}. Interaction
  \texttt{e} is implemented by step \texttt{6} (for FSL: through a
  modified version of the fsl\_sub script available in
  \url{https://github.com/aces/cbrain-plugins-neuro}; for PSOM:
  through a specific development in PSOM).}
\label{fig:cbrain-sub-tasking}
\end{figure*}

@@

The pool model was implemented in the SHIWA pool~\cite{ROGE-13}
diagrammed in Figure~\ref{fig:shiwa-pool-architecture}. In this
Figure, interaction \texttt{d} of Figure~\subref*{archi:agent} is
implemented in 3 different types of calls to the pool: workflow
submission (\texttt{1} \& \texttt{2}), workflow status retrieval
(\texttt{0} \& \texttt{3})\closednote{sorina}{I suppose letters come here
from the initial pool-model figure; it's a bit puzzling when mixed with
letters from Fig 2}{Tristan}{Fixed.}, and workflow retrieval (\texttt{13} \&
\texttt{14}). \note{Marc-e}{Maybe easier to follow if the following would be in fig
 legend instead of here, as for 2 previous figures} In Figure~\ref{fig:shiwa-pool-architecture},
interaction \texttt{f} of Figure~\subref*{archi:agent} is implemented
through 2 types of calls: workflow instance retrieval (\texttt{4},
\texttt{5} and \texttt{6}), and workflow instance update (\texttt{11}
and \texttt{12}). Workflow instance retrieval is used by the agents to
fetch work from the pool. Workflow instance update is used by the
agents to update workflow statuses. In the SHIWA pool, agents can wrap
different types of workflow engines to execute workflows expressed
with different languages. Calls \texttt{0}, \texttt{7}, \texttt{8} and
\texttt{9} on Figure~\ref{fig:shiwa-pool-architecture} are used by
workflow engine plugins to declare their supported language and launch
engines, and by workflow engines to report their status to the
agent. These calls are specific to the SHIWA Pool implementation of
the \texttt{agent} component and therefore have no corresponding
representation in Figure~\subref*{archi:agent}.

\begin{figure*}
\centering
\includegraphics[width=1.5\columnwidth]{figures/pool-interactions.pdf}
\caption{Architecture of the SHIWA pool. Circle-terminated arrows indicate messages
  that are broadcast to all pool clients. Figure reproduced
  from~\cite{ROGE-13}.}
\label{fig:shiwa-pool-architecture}
\end{figure*}

@@

Nested workflows have long been available in workflow engines, for
instance in the Taverna workbench~\cite{oinn2004taverna}. They are
also used implicitly in several platforms where workflow engines are
wrapped in workflow tasks as any other command-line tool. Nested
workflows were notably used by the SHIWA Science Gateway to implement
so-called Coarse-Grained workflow
interoperability~\cite{terstyanszky2014enabling}, i.e. to integrate
various workflow engines in a consistent
platform. Figure~\ref{fig:shiwa-architecture} shows the architecture
used in the SHIWA Simulation Platform for nested workflow execution
with service invocation.
\begin{figure*}
\centering
\includegraphics[width=1.5\columnwidth]{figures/shiwa-science-gateway.pdf}
\caption{Nested workflow execution through SHIWA Science Gateway. The
  parent workflow engine is \texttt{WS-PGRADE}, invoked as a service
  in the Science Gateway (step \texttt{1}, interaction \texttt{d} on
  Figure 2(V)-Right. Ten different child engines can be used by nested
  workflows, invoked through the \texttt{Submission service} (step
  \texttt{2}, interaction \texttt{d}). Each of these engines can
  submit tasks and transfer data to a distributed computing
  infrastructure (\texttt{DCI}, step \texttt{3}, interactions
  \texttt{b} and \texttt{c$_2$}). Data interactions (\texttt{c}) and
  application porting ones (\texttt{a}) are not represented. Figure
  reproduced from~\cite{terstyanszky2014enabling} with permission of
  the first author.}
\label{fig:shiwa-architecture}
\end{figure*}

@@

\bibliographystyle{elsarticle-num}
\bibliography{biblio}

>>>>

\begin{figure*}
\centering
\def\svgwidth{1.5\columnwidth}
\input{figures/VIP.pdf_tex}
\caption{Architecture of the Virtual Imaging Platform (service
  invocation).  Step \texttt{3} corresponds to interaction \texttt{d}
  on Figure~\ref{fig:architectures}(II), step \texttt{2} corresponds to
  interaction \texttt{c$_1$}, and step \texttt{4} maps to interactions
  \texttt{b} and \texttt{c$_2$} (in VIP, workflow data transfers are
  performed by the tasks and embedded in their descriptions).}
\label{fig:vip-architecture}
\end{figure*}

@@

\subsection{Task encapsulation}
\closednote{Marc-e}{May I propose
  another name here? Either "Task-centric" or "Task Oriented". My
  reasoning is that the WE is just another task, as well as the
  sub-tasks themselves.}{Tristan}{I agree: sub-tasking is not the
  integration model, it is the process that engines use when they
  are integrated as tasks. I think 'task encapsulation' is more
  consistent  with 'service invocation' than 'task-oriented'}

@@

\begin{figure*}
\centering
\subfloat[CBRAIN-FSL integration.]{
\def\svgwidth{\columnwidth}
\input{figures/CBRAIN-FSL.pdf_tex}
\label{fig:cbrain-fsl-architecture}
} \hfill \subfloat[CBRAIN-PSOM integration. Figure
reproduced from~\cite{GLAT-16}]{ \def\svgwidth{\columnwidth}
  \input{figures/CBRAIN-PSOM.pdf_tex}
\label{fig:cbrain-psom-architecture}
}
\caption{CBRAIN architecture for workflow engine integration
  (task encapsulation).  1: User sends data and workflow execution request to
  storage server(s) and CBRAIN portal. 2: CBRAIN portal sends
  execution request to execution server on cluster. 3: Execution
  server transfers data from storage server(s). 4-5: Execution server
  starts workflow engine (FSL tool or PSOM master) via task
  scheduler. 6: Workflow engine submits sub-tasks to execution server
  (FSL tasks or PSOM agents). 7-8: Execution server starts sub-tasks
  through task scheduler. FSL sub-tasks will run locally
  instead of being submitted to CBRAIN through interaction 6. 8':
  (PSOM only) PSOM master and workers execute workflow. 9. Execution
  server transfers results to storage server(s). Interaction
  \texttt{b} in Figure~\ref{fig:architectures}(III) is implemented by steps
  \texttt{4}, \texttt{5}, \texttt{7} and \texttt{8} (regular
  interactions with batch manager). Interaction \texttt{c} is
  implemented by steps \texttt{3} and \texttt{9}. Interaction
  \texttt{e} is implemented by step \texttt{6} (for FSL: through a
  modified version of the fsl\_sub script available in
  \url{https://github.com/aces/cbrain-plugins-neuro}; for PSOM:
  through a specific development in PSOM).}
\label{fig:cbrain-sub-tasking}
\end{figure*}

@@

The pool model was implemented in the SHIWA pool~\cite{ROGE-13}
diagrammed in Figure~\ref{fig:shiwa-pool-architecture}. In the SHIWA
pool, agents can wrap different types of workflow engines to execute
workflows expressed with different languages.

\begin{figure*}
\centering
\includegraphics[width=1.5\columnwidth]{figures/pool-interactions.pdf}
\caption{Architecture of the SHIWA pool. Circle-terminated arrows
  indicate messages that are broadcast to all pool
  clients. Interaction \texttt{d} of Figure~\ref{fig:architectures}(IV) is
  implemented in 3 different types of calls to the pool: workflow
  submission (\texttt{1} \& \texttt{2}), workflow status retrieval
  (\texttt{0} \& \texttt{3})\closednote{sorina}{I suppose letters come
    here from the initial pool-model figure; it's a bit puzzling when
    mixed with letters from Fig 2}{Tristan}{Fixed.}, and workflow
  retrieval (\texttt{13} \& \texttt{14}). \closednote{Marc-e}{Maybe
    easier to follow if the following would be in fig legend instead
    of here, as for 2 previous figures}{Tristan}{fixed} Interaction
  \texttt{f} is implemented through 2 types of calls: workflow
  instance retrieval (\texttt{4}, \texttt{5} and \texttt{6}), and
  workflow instance update (\texttt{11} and \texttt{12}). Workflow
  instance retrieval is used by the agents to fetch work from the
  pool. Workflow instance update is used by the agents to update
  workflow statuses.  Calls \texttt{0}, \texttt{7}, \texttt{8} and
  \texttt{9} are used by
  workflow engine plugins to declare their supported language and to
  launch engines, and by workflow engines to report their status to
  the agent. These calls are specific to the SHIWA Pool implementation
  of the \texttt{agent} component and therefore have no corresponding
  representation in Figure~\ref{fig:architectures}(IV). Figure reproduced
  from~\cite{ROGE-13}.}
\label{fig:shiwa-pool-architecture}
\end{figure*}

@@

Nested workflows have long been available in workflow engines, for
instance in the Taverna workbench~\cite{oinn2004taverna}. They are
also used implicitly in several platforms where workflow engines are
wrapped in workflow tasks as any other command-line tool. Nested
workflows were notably used by the SHIWA Science Gateway to implement
so-called Coarse-Grained workflow
interoperability~\cite{terstyanszky2014enabling}, i.e. to integrate
various workflow engines in a consistent
platform. Figure~\ref{fig:shiwa-architecture} shows the architecture
used in the SHIWA Simulation Platform for nested workflow execution
with service invocation.
\begin{figure*}
\centering
\includegraphics[width=1.5\columnwidth]{figures/shiwa-science-gateway.pdf}
\caption{Nested workflow execution through SHIWA Science Gateway. The
  parent workflow engine is \texttt{WS-PGRADE}, invoked as a service
  in the Science Gateway (step \texttt{1}, interaction \texttt{d} on
  Figure~\ref{fig:architectures}(V)-Right. Ten different child engines can be used by nested
  workflows, invoked through the \texttt{Submission service} (step
  \texttt{2}, interaction \texttt{d}). Each of these engines can
  submit tasks and transfer data to a distributed computing
  infrastructure (\texttt{DCI}, step \texttt{3}, interactions
  \texttt{b} and \texttt{c$_2$}). Data interactions (\texttt{c}) and
  application porting ones (\texttt{a}) are not represented. Figure
  reproduced from~\cite{terstyanszky2014enabling} with permission of
  the first author.}
\label{fig:shiwa-architecture}
\end{figure*}

@@

\enlargethispage{5mm}
\bibliographystyle{elsarticle-num}
\bibliography{biblio}
